On the presentation of orthogonal groups over polynomial rings

### 2019-01-28 dl-paper.tex
Generalized Proposition 1

\theoremstyle{definition}
\newtheorem{df}[lemma]{Definition} \Crefname{df}{Definition}{Definitions}
\newtheorem{example}[lemma]{Example} \Crefname{example}{Example}{Examples}
\newtheorem{rem}[lemma]{Remark}
\newtheorem{cond}[lemma]{Condition}
\DeclareMathOperator{\Ker}{Ker}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\G}{G}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\HH}{H}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\GW}{GW}
\DeclareMathOperator{\colim}{colim}
\newcommand{\XX}{\mathcal{X}}           % (subset of) Steinberg generators
\newcommand{\RR}[1]{\mathcal{R}_{#1}}   % Steinberg relations

@@

In order to better understand $C(\Phi, R, I)$  we would like to continue the sequence~\eqref{eq:relative-St} to the left. Thus we need to work with unstable $\K_3$-functors a la Quillen (cf.~\cite{Pa89}).
\begin{df}  
Denote by $X$ the topological space obtained from the (discrete) classifying space $B\G(\Phi, R)$ by means of Quillen's $+$-construction with respect to the elementary subgroup $\E(\Phi, R)$ (which is perfect thanks to the first assumption). Now set $\K_3(\Phi, R) = \pi_3(X)$. \end{df}
%TODO: Give reference on the definition of the plus-construction?

\begin{prop}\label{characterization} The group $\K_3(\Phi, R)$ is isomorphic to $\HH_3(\St(\Phi, R), \ZZ)$. In particular, it does not depend on the isogeny form of $\G(\Phi, R)$ used to define it. Moreover, for a ring epimorphism $R \to R/I$ there is a functorial exact sequence
\[ \xymatrix{\K_3(\Phi, R) \ar[r] & \K_3(\Phi, R/I) \ar@{->>}[r] & C(\Phi, R, I).} \]
\end{prop}
\begin{proof}
 %TODO: First part.

@@

\begin{lemma} \label{lem:Panin-stability}
 Let $R$ be either a field, principal ideal domain or a Dedekind domain. 
 Then the canonical map $\K_3(\rD_\ell, R) \to \K_3(\rD_\infty, R)$ is an epimorphism for $\ell \geq 6$ 
 and an isomorphism for $\ell \geq 7$.
\end{lemma}
\begin{proof}
 Stabilization for orthogonal and symplectic K-groups a la Quillen is given by Panin's theorem, see~\cite[Theorem~9.4]{Pa89}. More specifically the bounds for surjective (resp. injective) stability is 
 $\ell \geq \max(6, a+2)$ (resp. $\ell \geq \max(7,a+3)$) for $a=1,2,3$ in the case when $R$ is a field, p.i.d. or a Dedekind domain, respectively.
\end{proof}

@@

\begin{lemma} Let $\Phi$ be a reduced irreducible root system of type $\neq \rG_2$ and $k$ be arbitrary field.
Then there is a split exact sequence of abelian groups
\[ \xymatrix{ 0 \ar[r] & \K_2(\Phi, k) \ar[r] & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & H(\Phi, k) \ar[r] & 0,}\text{ where} \]
\[ H(\Phi, k) = \left\{\begin{array}{ll} \K_1^\mathrm{MW}(k)& \text{if $\Phi$ is symplectic,}\\ k^* & \text{otherwise.}  \end{array}\right \]  \end{lemma}
\begin{proof} In the nonsymplectic case one can choose a long simple root $\alpha$ of $\Phi$ in such a way that there is a commutative diagram of abelian groups
\[\xymatrixcolsep{5pc}\xymatrix{k^* \ar[r]^-{\{ -, X \}_{\alpha}} \ar[rrd]_{\{-, X\}} & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & \K_2(\Phi, k(X)) \ar[d]^{\cong} \\
                                                                                      &                                 & \K_2^\mathrm{M}(k(X)).} \]
Clearly, the diagonal arrow is injective by Milnor's theorem and the vertical arrow is an isomorphism by Matsumoto's theorem~\cite[Theorem~5.10]{Ma69}.
The required assertion now follow from~\cite[Satz~3]{Hur77} and the fact that the left horizontal arrow is injective.

>>>>

\theoremstyle{definition}
\newtheorem{df}[lemma]{Definition} \Crefname{df}{Definition}{Definitions}
\newtheorem{example}[lemma]{Example} \Crefname{example}{Example}{Examples}
\newtheorem{rem}[lemma]{Remark}
\newtheorem{cond}[lemma]{Condition}

\newlist{proplist}{enumerate}{1} \setlist[proplist]{label=(\roman{proplisti}), ref=\thethm.(\roman{proplisti}),noitemsep} \Crefname{proplisti}{Proposition}{Propositions}
\newlist{lemlist}{enumerate}{1} \setlist[lemlist]{label=(\roman{lemlisti}), ref=\thelem.(\roman{lemlisti}),noitemsep} \Crefname{lemlisti}{Lemma}{Lemmas}

\DeclareMathOperator{\Ker}{Ker}
\DeclareMathOperator{\St}{St}
\DeclareMathOperator{\G}{G}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\HH}{H}
\DeclareMathOperator{\K}{K}
\DeclareMathOperator{\GW}{GW}
\DeclareMathOperator{\colim}{colim}
\newcommand{\XX}{\mathcal{X}}           % (subset of) Steinberg generators
\newcommand{\RR}[1]{\mathcal{R}_{#1}}   % Steinberg relations
\newcommand{\catname}[1]{{\normalfont\textbf{#1}}} %Category name

@@

\subsection{Unstable Quillen K-groups}
In order to better understand $C(\Phi, R, I)$  we would like to continue the sequence~\eqref{eq:relative-St} to the left. 
Thus we need to work with unstable $\K_3$-functors a la Quillen (cf.~\cite{Pa89}).

Let $G \colon \catname{CRings} \to \catname{Groups}$ be a functor and $P$ its subfunctor such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$.

Thanks to the functoriality of Quillen's plus-construction (see e.\,g.~\cite[Proposition~5.2.4]{Ro95}) one can define a functor 
$X_{G, P}(R) = BG(R)^+_{P(R)} \colon \catname{CRings} \to \catname{Top}$ and then define the corresponding K-group functors via $\K_{i}(G, P, R) = \pi_i(X_{G, P}(R))$.

\begin{prop}\label{characterization}
Let $S \colon \catname{CRings} \to \catname{Groups}$ be another group functor such that
 for all $R$ the group $S(R)$ is a universal central extension of $P(R)$. Then the following holds:
\begin{proplist}
 \item There is a natural isomorphism $\K_3(G, P, R) \cong \HH_3(S(R), \ZZ)$.
 \item For a ring epimorphism $R \to R/I$ there is a natural exact sequence
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & \HH_3(S(R), S(R/I), \ZZ).} \]
\end{proplist}
\end{prop}
\begin{proof}
 Follows from the solution of Exercises~1.8--1.9 of \cite[Chapter~IV]{Kbook} provided one notices that all the intermediate isomorphisms are natural in $R$.

@@

Stability for orthogonal and symplectic K-theories is given by the following theorem of Panin
\begin{theorem}[Panin] \label{lem:Panin-stability}
 Let $R$ be either a field, principal ideal domain or a Dedekind domain and set $a = 1,2,3$ in each of respective cases.
 Then the canonical map $\K_k(\rD_\ell, R) \to \K_k(\rD_\infty, R)$ is an epimorphism for $\ell \geq \mathrm{max}(2k, a+k-1)$ 
 and an isomorphism for $\ell \geq \mathrm{max}(2k+1, a+k)$, where $a=1,2,3$ in the case when $R$ is a field, p.i.d. or a Dedekind domain, respectively \end{theorem}
\begin{proof} See~\cite[Theorem~9.4]{Pa89}. \end{proof}

\begin{comment}
 Stabilization for orthogonal and symplectic K-groups a la Quillen is given by Panin's theorem,  More specifically the bounds for surjective (resp. injective) stability is 
 $\ell \geq \max(6, a+2)$ (resp. $\ell \geq \max(7,a+3)$) for $a=1,2,3$ in the case when $R$ is a field, p.i.d. or a Dedekind domain, respectively.
\end{comment}

@@

\begin{lemma} Let $\Phi$ be a reduced irreducible root system of type $\neq \rG_2$ and $k$ be arbitrary field.
Then there is a split exact sequence of abelian groups
\[ \xymatrix{ 0 \ar[r] & \K_2(\Phi, k) \ar[r] & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & H(\Phi, k) \ar[r] & 0,}\text{ where} \]
\[ H(\Phi, k) = \left\{\begin{array}{ll} \K_1^\mathrm{MW}(k)& \text{if $\Phi$ is symplectic,}\\ k^* & \text{otherwise.}  \end{array}\right. \]  \end{lemma}
\begin{proof} In the nonsymplectic case one can choose a long simple root $\alpha$ of $\Phi$ in such a way that there is a commutative diagram of abelian groups
\[\xymatrixcolsep{5pc}\xymatrix{k^* \ar[r]^-{\{ -, X \}_{\alpha}} \ar[rrd]_{\{-, X\}} & \K_2(\Phi, k[X, X^{-1}]) \ar[r] & \K_2(\Phi, k(X)) \ar[d]^{\cong} \\
                                                                                      &                                 & \K_2^\mathrm{M}(k(X)).} \]
Clearly, the diagonal arrow is injective by Milnor's theorem and the vertical arrow is an isomorphism by Matsumoto's theorem~\cite[Theorem~5.10]{Ma69}.
The required assertion now follow from~\cite[Satz~3]{Hur77} and the fact that the left horizontal arrow is injective.
